The semantic drift of quotations in blogspace: a case study in short-term cultural evolution

### 2016-07-18 brains-protocol.tex
fix/protocol: fixes after Camilles comments

Quotations appeared to be a perfect candidate to propose a first measure of \newtext{automatic} cognitive bias in \newtext{cultural transmission}.
First, they are usually cleanly delimited by quotation marks which greatly facilitates their detection in text corpora.
Second, they stem from a unique original version, and are ideally traceable back to that version.
Third, and most importantly, their duplication should \emph{a priori} be highly faithful, apart from cases of cropping:
not only should transformations be of moderate magnitude, but when specific words are not perfectly duplicated, it is safe to assume that the variation is due to involuntary cognitive bias --- as writers may expect any casual reader to easily verify, and thus criticize, the fidelity to the original quotation.

We could therefore study the individual transformation process at work when authors alter quotations, by examining the modified words in each transformation.
\newtext{Since our approach is exploratory however, we do not know at the outset which precise effect of cognitive bias we are looking for.
Indeed the data we use does not come from a controlled experiment in the laboratory, designed to elicit a particular effect: they are recordings of real life interactions, with all the complexity and uncertainty of conditions this entails.
In this first study, therefore, we did not try to predict and explain in detail the cognitive processes responsible for transformations, as this would be akin to a drink from the firehose.
Our goal, instead, is to show that such effects exist and are measurable, even if they are part of a larger complexity.
That is, we aim to prove first a relatively weak hypothesis before trying stronger ones.
If this is confirmed, we will have successfully applied laboratory analyses to out-of-laboratory data, opening a path to explanations of actual (\hbox{vs.} simulated) cultural evolution with tools from cognitive science.
As explained in the previous section, this is the reason we chose to use measures that can aggregate over all the transformations in the data set.}

@@

We used a quotation data set collected by \citet{leskovec_meme-tracking_2009}, large enough to lend itself to statistical analysis.
This data set consists of the daily crawling of news stories and blog posts from around a million online sources, with an approximate publication rate of 900k texts per day, over a nine-month period of time from August 2008 to April 2009 \newtext{(\citealp{leskovec_memetracker:_2009}; unfortunately, the original article, \citealp{leskovec_meme-tracking_2009}, does not provide further details on the source selection methodology).}
\newtext{The authors automatically extracted quotations} from this corpus.
Each quotation is a more or less faithful excerpt of an utterance (oral or written) by the quoted person; for instance:
\begin{quote}
The Bank of England said, "these operations are designed to address funding pressures over quarter-end."
\end{quote}

@@

Since our data is about written transformations, number of letters and orthographic neighborhood density are the natural representatives of groups (a) and (b) respectively.
Given the importance of age of acquisition in the lexical feature literature, we chose it to represent group (c).
Finally we used clustering coefficient to represent group (d) since it has already been used in previous studies.
The final set of features we will discuss in the rest of the paper, as well as their cross-correlations, can be seen in Fig.~\ref{fig:feature-corrs-filtered} (the analysis on the complete set of features can be found in Supplementary Material).\footnote{
Note that feature values stem from different data sets which do not always encode the same words.
Indeed, we have data on frequency for about \newtext{33.5k} words, on age of acquisition for 30.1k words, on clustering coefficient for 5.7k words, number of synonyms 111.2k, and \newtext{orthographic density 17.8k words}.
Quite often then, not all features are available for all words in our data set;
however this is not problematic since the analysis is done on a per-feature basis, and not all words need be encoded in all features.}

>>>>

\newtext{We rely on a text corpus made of quotations extracted from online blog posts, and focus on their evolution.}
\newtext{Indeed} quotations appeared to be a perfect candidate to propose a first measure of \newtext{automatic} cognitive bias in \newtext{cultural transmission}.
First, they are usually cleanly delimited by quotation marks which greatly facilitates their detection in text corpora.
Second, they stem from a unique original version, and are ideally traceable back to that version.
Third, and most importantly, their duplication should \emph{a priori} be highly faithful, apart from cases of cropping:
not only should transformations be of moderate magnitude, but when specific words are not perfectly duplicated, it is safe to assume that the variation is due to involuntary cognitive bias --- as writers may expect any casual reader to easily verify, and thus criticize, the fidelity to the original quotation.

We could therefore study the individual transformation process at work when authors alter quotations, by examining the modified words in each transformation.
\newtext{Since our approach is exploratory however, we do not know at the outset which precise effect of cognitive bias we are looking for.
Indeed the data we use does not come from a controlled experiment in the laboratory, designed to elicit a particular effect: they are recordings of real life interactions, with all the complexity and uncertainty of conditions this entails.
%In this study, therefore, we did not try to predict and explain in detail the cognitive processes responsible for transformations
%\todo{, as this would be akin to a drink from the firehose. \sf [Cam: supprimer? potentiellement un excès d'autoflagellation -- en fait on devrait pouvoir dire simplement: ``The prediction and detailed explanation of the cognitive processes responsible for transformations is outside of the scope of this study'', probablement après la fin de la phrase suivante (``of a larger complexity (the detailed prediction and deconstruction of the cognitive processes etc. outside the scope etc. further research)'' mais même là cette modestie pourrait plutôt être de mise dans la conclusion? ou bien il s'agit d'une critique directe d'un reviewer ?  Un peu pareil pour ``our approach is exploratory'' (que j'ai laissé au nom de l'ambiguïté d'exploratoire qui veut aussi bien dire ``premier brouillon'' que ``data mining'') et ``this first study'' (où j'ai enlevé le `first' car trop ``bon c'est à moitié fini mais on soumet quand même'' ;) ]}
Our goal, therefore, is to show that such effects exist and are measurable even if they are part of a larger complexity (the detailed prediction and deconstruction of the cognitive processes responsible for them being left to further research).
%\todo{That is, we aim to prove first a relatively weak hypothesis before trying stronger ones. [Cam: idem, je supprimerais celle-là]}
If this is confirmed, we will have successfully applied laboratory analyses to out-of-laboratory data, opening a path to explanations of actual (\hbox{vs.} simulated) cultural evolution with tools from cognitive science.
As explained in the previous section, this is the reason we chose to use measures that can aggregate over all the transformations in the data set.}

@@

We used a quotation data set collected by \citet{leskovec_meme-tracking_2009}, large enough to lend itself to statistical analysis.
This data set consists of the daily crawling of news stories and blog posts from around a million online sources, with an approximate publication rate of 900k texts per day, over a nine-month period of time from August 2008 to April 2009 (\citealp{leskovec_memetracker:_2009}).\footnote{
The original article~\citep{leskovec_meme-tracking_2009} does not provide further details on the source selection methodology.
}
\newtext{The authors automatically extracted quotations} from this corpus.
Each quotation is a more or less faithful excerpt of an utterance (oral or written) by the quoted person; for instance:
\begin{quote}
The Bank of England said, "these operations are designed to address funding pressures over quarter-end."
\end{quote}

@@

Since our data is about written transformations, number of letters and orthographic neighborhood density are the natural representatives of groups (a) and (b) respectively.
Given the importance of age of acquisition in the lexical feature literature, we chose it to represent group (c).
Finally we used clustering coefficient to represent group (d) since it has already been used in previous studies.
The final set of features we will discuss in the rest of the paper, as well as their cross-correlations, can be seen in Fig.~\ref{fig:feature-corrs-filtered} (the analysis on the complete set of features can be found in the appendix).\footnote{
Note that feature values stem from different data sets which do not always encode the same words.
Indeed, we have data on frequency for about \newtext{33.5k} words, on age of acquisition for 30.1k words, on clustering coefficient for 5.7k words, number of synonyms 111.2k, and \newtext{orthographic density 17.8k words}.
Quite often then, not all features are available for all words in our data set;
however this is not problematic since the analysis is done on a per-feature basis, and not all words need be encoded in all features.}
